---
title: "ESM244_lab1"
author: "Sam Csik"
date: "1/10/2019"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Objectives:

-basic review (data wrangling)
-intro to spatial data in R with sf
-intro to shiny apps

Part 1. Attach packages
```{r}
library(tidyverse)
library(sf)
library(leaflet)
library(tmap)
```

Part 2. Review on reading, exploring, wrangling, and visualizing data

a. Read in the data
```{r}
wine <- read_csv("wine_data.csv")
```

b. Wrangle data
```{r}
wine_new <- wine %>% 
  select(country, province, winery, region_1, points, price) %>% 
  rename(state = province) %>% 
  filter(state == "California" | state == "Oregon" | state == "Washington") %>% 
  mutate(ppd = points/price) %>% # add new columns
  arrange(-ppd) # arrange in decreasing order of points per dollar
```

c. Use group_by() and summarize() to find summary statistics by group
```{r}
wine_summary <- wine_new %>% 
  group_by(state) %>% 
  summarise(
    mean_ppd = mean(ppd, na.rm = TRUE),
    max_ppd = max(ppd, na.rm = TRUE)
  )
```

d. A few gg-graphics
```{r}
wine_hist <- ggplot(wine_new, aes(x = ppd)) + 
  geom_histogram(aes(fill = state)) +
  theme_classic() +
  theme(axis.text = element_text(color = "black"), 
          panel.border = element_rect(colour = "black", fill=NA, size=0.7), 
          legend.position = "none") +
  scale_x_continuous(expand = c(0,0)) +
  scale_y_continuous(expand = c(0,0)) +
  facet_wrap(~state, scales = "free")
wine_hist
```

new graph: scatterplot grpah of price vs. points, with color determined by state, adding a linear model trendline on top of it
```{r}
wine_points <- ggplot(wine_new, aes(x = price, y = points)) +
  geom_point(aes(color = state), alpha = 0.5) +
  geom_smooth(color = "black", method = "lm", se = FALSE) +
  scale_color_manual(values = c("royalblue", "mediumorchid1", "orange")) +
  facet_wrap(~state) +
  theme_classic() +
  theme(axis.text = element_text(color = "black"), 
          panel.border = element_rect(colour = "black", fill=NA, size=0.7), 
          legend.position = "none") 
  
wine_points 
```

Part 3. California counties (spatial data) using sf
```{r}
ca_counties <- st_read(dsn = ".", layer = "CA_Counties_TIGER2016") # "." means we're already in our wd; layer = "first common string name of each file used"

ca_land <- ca_counties %>% 
  select(NAME, ALAND)

ca_pop_inc <- read_csv("ca_pop_inc.csv")  %>% 
  rename(NAME = COUNTY) 

ca_df <- full_join(ca_land, ca_pop_inc) %>% # merge ca_pop_inc to ca_land; full_join will keep every row, even if there isn't a match
  select(NAME, MedFamilyIncome)

# plot(ca_counties)

```

d. use geom_sf to make a map in ggplot
```{r}
ca_income <- ggplot(ca_df) +
  geom_sf(aes(fill = MedFamilyIncome))

ca_income

ca_df_transform <- st_transform(ca_df, crs = 4326)

leaflet(ca_df_transform) %>% 
  addTiles() %>% 
  addPolygons()
```


